A Fast Unified Model for Parsing and Sentence Understanding

### 2016-03-18 writing/hard_stack_paper/paper.tex
More section 3 edits

Lexical ambiguity is ubiquitous in natural language---most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

We propose an alternative space-efficient stack representation technique inspired by the zipper \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

@@

This stack representation requires substantially less space. It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation. Furthermore, all of the updates to $S$ and $Q$ can be performed in-place on a GPU. These features allow this stack model to run efficiently on a GPU. We describe speed results in Section~\ref{sec:speed}.

>>>>

Lexical ambiguity is ubiquitous in natural language. Most words have multiple senses or meanings, and it is generally necessary to use the context in which a word occurs to determine which of its senses or meanings is meant in a given sentence. Simpler sequence-based sentence encoding models like the standard LSTM RNN have an advantage here: when a sequence-based model first processes a word, it has direct access to a state vector that summarizes the left context of that word, which acts as a cue for disambiguation. In contrast, when standard tree-structured models first process a word, they only have access to the constituent that the word is merging with, which is often just a single additional word. Feeding a context representation from the tracking LSTM into the composition function is a simple and efficient way to mitigate this disadvantage of tree-structured models.

@@

We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}. For a single input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. We maintain a queue of backpointers onto $S$ in order to track which elements should be involved in a \reduce~operation at any given time. Algorithm~\ref{alg:thin-stack} below describes the full mechanics of a stack feedforward in this compressed representation. It operates on the compressed $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run of this algorithm.

@@

This stack representation requires substantially less space. It stores each element involved in the feedforward computation exactly once, meaning that this representation can still support efficient backpropagation. Furthermore, all of the updates to $S$ and $Q$ can be performed batched and in-place on a GPU, yielding substantial speed gains. We describe speed results in Section~\ref{sec:speed}.
